---
title: "ST558 Project 2"
author: "Sergio Mora & Ashley Ko"
output: html_document
params:
  category: "lifestyle"
---



```{r setup, include=FALSE, message=FALSE}
set.seed(123)

#Required Packages
library(knitr)
library(rmarkdown)
library(caret)
library(tidyverse)
library(dplyr)
```  

# Introduction

# Data
## Reading in data
We will read in our csv dataset. As instructed we will also split our data by `data_channel_is_*`.
```{r}
online_new_popularity_data <- read.csv("./OnlineNewsPopularity/OnlineNewsPopularity.csv")
```

## Subsetting the data
We will subset the data based on the category listed in our YAML. In this case we will be using ``r paste0("data_channel_is_",params$category)``.
```{r}
#Subsetting our data based on the category parameter
subset_data <- online_new_popularity_data %>% filter(!!as.name(paste0("data_channel_is_",params$category)) == 1) %>% select(-starts_with("data_channel_is"))

#Setting up a simple 70/30 split for our already subset data
sample_size <- floor(0.7 * nrow(subset_data)) 
train_ind <- sample(seq_len(nrow(subset_data)), size = sample_size)

# This will be needed later on when we start modeling
training_data <- subset_data[train_ind,]
test_data <- subset_data[-train_ind,]

```

# Summarizations
## EDA's

Now that we have our data the way we want it we will perform some EDA's to be able to get a good grasp of what our data looks like.

```{r}
summary(subset_data$shares)
```



## Contigency tables

# Modeling

# Comparison